---
title: "Survey Map"
output:
  html_document: default
  pdf_document: default
---


```{r setup, include=FALSE}
library(readxl)
library(sf)
library(stars)
library(gstat)
library(viridis)
library(RColorBrewer)
library(ggmap)
library(tidyverse)
library(mapview)
library(classInt)
library(metR)
library(OpenStreetMap) 
library(svDialogs)
```



1. Definition of the parameters, units, detector, file name, …

```{r definitions}
rm(list = ls())
# Global variables

# OPENS WINDOWS if TRUE
OPEN_WINDOWS = FALSE

crs <- 4326

source_name <- c(
    "inputs/2023-10-20-Seibersdorf-B_PGIS-89BE_2023_10_10_14_03_37-dose.R"
    #"inputs/2023-11-21-Granite-mine-Medusa-processed-U.R"
    #,
    #"inputs/2023-11-21-Granite-mine-PGIS-09_55_36-U.R"
    #"inputs/2024-05-15-Seibersdorf-Arktis-cps.R"
    #"inputs/2024-10-16-Trieste-Gamon2.R"
    )
```

2.  Read the file

```{r readFile}
source(source_name[1])
if(endsWith(input$filename, ".xlsx")) {data <- read_excel(paste0("data/",input$filename))}
if(endsWith(input$filename, ".csv")) {data <- read.csv(paste0("data/",input$filename))}
 
  data <- data %>%   
    rename(lon = input$longitude, lat = input$latitude, value = input$column) %>% 
    select("lon","lat","value")

data$value <- data$value * input$constant
data$dataset <- input$detector
detectorname <- input$detector

```

3.  Data from more datasets (detectors)

```{r moreFiles}
if (length(source_name) > 1){

 for (src in 2:length(source_name)) {

  source(source_name[src])
  if(endsWith(input$filename, ".xlsx")) {data_temp <- read_excel(paste0("data/",input$filename))}
  if(endsWith(input$filename, ".csv")) {data_temp <- read.csv(paste0("data/",input$filename))}
  
  data_temp <- data_temp %>%   
   rename(lon = input$longitude, lat = input$latitude, value = input$column) %>% 
   select("lon","lat","value")
  data_temp$value <- data_temp$value * input$constant
  data_temp$dataset <- input$detector
  data <- rbind(data,data_temp)  # repeated with each dataset

  # for the detector name in the printouts
  detectorname <- paste0(detectorname,"+",input$detector) 
 
  #head(data)
  #tail(data)
 }

}

head(data)
tail(data)
summary(data, digits=6)
data %>%  count(dataset)
```

```{r moreFiles}
if (length(source_name) > 1){
 ggplot(data, mapping = aes(x = value)) +
  geom_histogram(fill = "steelblue") +
  facet_grid(dataset ~  .) 
}
```  
```{r moreFiles}  
if (length(source_name) > 1){
  ggplot(data, mapping = aes(x = lon, y = lat)) +
  geom_point(aes(color = value), alpha = .8, size = 1) +
  scale_color_viridis(direction = 1) +
  coord_quickmap() +
  facet_grid(dataset ~  .)
}
```

4.  Process sorted data

```{r sort}
#print the beginning and end of the data
head(data)
tail(data)

#space between white lines later in the maps
signif(((data %>% pull(value) %>% max()) - (data %>% pull(value) %>% min()))/2,1) -> ISObinwidth 

# Subsetting a dataframe by index of rows to be selected
# data[index_of_rows, columns_of_interest] (here we want to keep all columns)
data_sub <- data[sample(1:nrow(data), (nrow(data)/input$subsetting)),]

# Convert dataframe to an "sf" compatible object
data_sf_all <- st_as_sf(data, coords = c("lon", "lat"), crs = crs)
data_sf_sub <- st_as_sf(data_sub, coords = c("lon", "lat"), crs = crs)

# Remove duplicated locations:
# Based on https://github.com/r-spatial/sf/issues/669
d = st_is_within_distance(data_sf_all, dist=0.04) #originally 0.01
dupl = unlist(mapply(function(x,y) x[x < y], d, seq_along(d)))

if(length(dupl))data_sf <- data_sf_all[-dupl, ] else data_sf <- data_sf_all

# Transform to UTM
data_utm <- st_transform(data_sf, crs = input$utm)

# Produce perimeters and bboxes
site_perimeter_sf <- st_convex_hull(st_union(data_sf))
site_perimeter_utm <- st_convex_hull(st_union(data_utm))
bbox_sf <- st_bbox(data_sf)
bbox_utm <- st_bbox(data_utm)

#plot(site_perimeter_sf)
#plot(site_perimeter_utm)

# Generate grids covering the bounding box and cropped by the site's perimeter.
bbox_sf %>%
  st_as_stars(dx = input$resolution, dy = input$resolution) %>% 
  st_crop(site_perimeter_sf) -> grd_sf

bbox_utm %>%
  st_as_stars(dx = input$resolution, dy = input$resolution) %>%
  st_crop(site_perimeter_utm) -> grd_utm
```

5.  Show and draw data

DATA: data - raw data data_sf - data in the format of the sf library
data_utm - data in the utm format of the sf library

Plotting tools: plot - data format: data, data_sf, data_utm ggplot -
data format: data, data_sf, data_utm mapview - data format: data_sf,
data_utm OpenStreetMap - data format: data

```{r showData}
st_crs(data_sf)
st_crs(data_utm)

# Data visualization with ggplot2
# https://ggplot2.tidyverse.org/         install.packages("tidyverse")  or  install.packages("ggplot2")

ggplot(data, mapping = aes(x = value)) +
  geom_histogram(fill = "steelblue")

ggplot(data_sub, mapping = aes(x = value)) +
  geom_histogram(fill = "steelblue")

# Bounding boxes
plot(bbox_sf)
plot(bbox_utm)
plot(site_perimeter_utm)
plot(grd_utm)

#   - "st_geometry(data_sf)" specifies we only want to map the geometry. 
plot(data)
plot(data_sf)
plot(data_utm)


ggplot(data = data, mapping = aes(x = lon, y = lat)) +
  geom_point(aes(color = value), alpha = .6, size = 2) +
  scale_color_viridis(direction = 1) +
  coord_quickmap()

ggplot(data = data_sf) +
  geom_sf(aes(color = value)) +
  scale_color_viridis(direction = 1) +
  coord_sf(crs = crs)

ggplot(data = data_utm) +
  geom_sf(aes(color = value)) +
  scale_color_viridis(direction = 1) +
  coord_sf(crs = crs)

# Create an interactive map with "mapview" funtion from the "mapview" package
# https://r-spatial.github.io/mapview/    install.packages("mapview")
#   - zcol: variable to be colour encoded
#   - map.types: base layer (here Esri satellite imagery)
#   - legend: do we want a legend or not
mapviewOptions(platform = "leaflet")

mapview(data_sf_sub, zcol = "value", map.types = c("Esri.WorldImagery", "CartoDB.Positron", "CartoDB.DarkMatter", "OpenStreetMap", "OpenTopoMap"), legend = TRUE)
mapview(data_sf, zcol = "value", map.types = c("Esri.WorldImagery", "CartoDB.Positron", "CartoDB.DarkMatter", "OpenStreetMap", "OpenTopoMap"), legend = TRUE)
mapview(data_utm, zcol = "value", map.types = c("Esri.WorldImagery", "CartoDB.Positron", "CartoDB.DarkMatter", "OpenStreetMap", "OpenTopoMap"), legend = TRUE)
```

6.  Print maps of the walks

```{r printSurvey}
my_osmmap <- openmap(c(bbox_sf$ymax+0.2*(bbox_sf$ymax-bbox_sf$ymin),
                    bbox_sf$xmin-0.2*(bbox_sf$xmax-bbox_sf$xmin)), 
                    c(bbox_sf$ymin-0.2*(bbox_sf$ymax-bbox_sf$ymin),
                    bbox_sf$xmax+0.2*(bbox_sf$xmax-bbox_sf$xmin)), 
                              zoom=18, type="bing", mergeTiles = TRUE)
# Adjust the zoom level if needed

# Re-project it to WGS84 Coordinate Reference System
my_osmmap <- openproj(my_osmmap)

autoplot.OpenStreetMap(my_osmmap) + geom_point(aes(x = lon, y = lat, color = value),  alpha = 0.5, size = 2, data = data_sub)  +
  scale_color_viridis_c(name=quantity, option = "viridis", alpha=0.4)  +
  labs(title = site,
       subtitle = paste(detectorname,"detector"),
       x="Longitude", y="Latitude") +
  theme(legend.position="right",axis.text.x=element_text(angle = -45, hjust = 0))

 ggsave(paste0("printouts/",input$site,"-",detectorname,"-",input$unitName,"-survey.jpg"), dpi=600)
 if (OPEN_WINDOWS){
  viewer <- getOption("viewer") 
  viewer(paste0("printouts/",input$site,"-",detectorname,"-",input$unitName,"-survey.jpg"), height = 10)
 }
  #ggsave(paste0("printouts/",site,"-",detector,"-",unitName,"-survey.pdf"), dpi=600)

# Define your threshold level (in our case in nSV/h)
#value_threshold <- 60
#autoplot.OpenStreetMap(my_osmmap) + geom_point(aes(x = lon, y = lat), colour = "red", alpha = 1, size = 2,
#             data = data[data$value > value_threshold,]) +
#  theme(legend.position="right",axis.text.x=element_text(angle = -45, hjust = 0))

 #ggsave("printouts/above-threshold.jpg", dpi=600)
 #ggsave("printouts/above-threshold.pdf", dpi=600)
```

7.  Interpolation

Kriging

```{r krige}
value_vgm <- variogram(value ~ 1, data_utm, cloud = FALSE)
                      #,boundaries = c(seq(2,10,1), seq(11, 50, 3))) #, seq(41, 400, 5)))

nugget <- value_vgm$gamma[1]                        # first variogram value
psill <- max(value_vgm$gamma) - value_vgm$gamma[1]  # Partial sill from thge maximum value
range <- value_vgm$dist[which.max(value_vgm$gamma)] # position of the maximum

# Eyeball the variogram and estimate the initial parameters
#nugget <- 2
#psill <- 25 # Partial sill
#range <- 10

# To see models available, execute "vgm()" in RStudio console: Nug, Exp, ...
# They are "canonical" class of functions to be fitted. 
# "kappa" parametrizes some models: for instance a Matern model with 
# kappa = 0.5 is the exponential model. In our case, we ask to find this parameter
# automatically.
# Finding the right function, parameters is more an "art" than a science. So try and interpret ...
v_model <- fit.variogram(value_vgm, model = vgm(
    model = "Sph",
    nugget = nugget,
    psill = psill,
    range = range,
    fit.kappa = TRUE
) )


jpeg(file=paste0("printouts/",input$site,"-",detectorname,"-",input$unitName,"-variogram.jpg"))
plot(value_vgm, model = v_model)
dev.off()

if (OPEN_WINDOWS){
 viewer(paste0("printouts/",input$site,"-",detectorname,"-",input$unitName,"-variogram.jpg"), height = 10)
}

#jpeg(file=paste0("printouts/",input$site,"-",detectorname,"-",input$unitName,"-variogram.jpg"))
#plot(value_vgm, model = v_model)
#dev.off()

# We pass in to the kriging:
#   - the value to predict
#   - the dataset
#   - the output data
#   - our model from variogram
#   - we want to use max 100 surrounded points (would be computationally demanding otherwise)
value_krig <- krige(value ~ 1, data_utm[, "value"], newdata = grd_utm, model = v_model, nmax = 100)

#uncertainty of kriging in %
value_krig$cov <- 100 * sqrt( value_krig$var1.var) / value_krig$var1.pred

print(v_model)
plot(value_vgm, model = v_model)
```

Inverse distance weighting

```{r idw}
# idw: is the inverse distance weighting power (the vicinity of points to be considered).
#      Experiment several values and see how smooth or sharp your map becomes.
value_idw <- idw(value ~1, data_utm, grd_utm, idp = 10)
value_idw 
```

8.  Show maps with the interpolated values

```{r showInterpolations, warning = FALSE, message = FALSE}

plot(value_krig["var1.pred"], breaks="equal", col = hcl.colors(11, "Spectral", rev = TRUE))
plot(value_krig["var1.var"], breaks="equal", col = hcl.colors(11, "Spectral", rev = TRUE))
plot(value_krig["cov"], breaks="equal", col = hcl.colors(11, "Spectral", rev = TRUE))

plot(value_idw["var1.pred"], breaks="equal", col = hcl.colors(11, "Spectral", rev = TRUE))

ggplot(data = as.data.frame(value_krig),
       aes(x = x, y = y, fill = var1.pred)) +
  geom_raster() +
  scale_fill_viridis_c(name=quantity, 
                       option = "viridis", na.value = "transparent") +
  geom_contour(aes(z=var1.pred), binwidth = ISObinwidth,
               colour ="white", size = 0.1)  +
  geom_text_contour(aes(z = var1.pred), binwidth = ISObinwidth, colour = "white", size=3) +
  coord_sf(expand = F) +
  labs(title = input$site,
       subtitle = paste(detectorname,"detector krig"),
       x="X", y="Y") +
  theme(legend.position="right")

ggplot(data = as.data.frame(value_krig),
       aes(x = x, y = y, fill = cov)) +
  geom_raster() +
  scale_fill_viridis_c(name="uncertainty [%]", 
                       option = "viridis", na.value = "transparent") +
  geom_contour(aes(z=var1.pred), binwidth = ISObinwidth,
               colour ="white", size = 0.1)  +
  geom_text_contour(aes(z = var1.pred), binwidth = ISObinwidth, colour = "white", size=3) +
  coord_sf(expand = F) +
  labs(title = site,
       subtitle = paste(detectorname,"detector krig"),
       x="X", y="Y") +
  theme(legend.position="right")

ggplot(data = as.data.frame(value_idw),
       aes(x = x, y = y, fill = var1.pred)) +
  geom_raster() +
  scale_fill_viridis_c(name=input$quantity, 
                       option = "viridis", na.value = "transparent") +
  geom_contour(aes(z=var1.pred), binwidth = ISObinwidth,
               colour ="white", size = 0.1)  +
  geom_text_contour(aes(z = var1.pred), binwidth = ISObinwidth, colour = "white", size=3) +
  coord_sf(expand = F) +
  labs(title = input$site,
       subtitle = paste(detectorname,"detector idw"),
       x="X", y="Y") +
  theme(legend.position="right")

mapview(value_krig, alpha.regions = 0.7, legend = TRUE, map.types = "Esri.WorldImagery", col = hcl.colors(11, "Spectral", rev = TRUE))
mapview(value_idw, alpha.regions = 0.7, legend = TRUE, map.types = "Esri.WorldImagery", col = hcl.colors(11, "Spectral", rev = TRUE))
```

9.  Print the maps

```{r printMaps}
# re-project and generate a regular grid based on the bounding box
value_krig %>% st_transform(crs = crs) %>% st_bbox() %>% 
  st_as_stars() -> new_grd

# Warp it
# For further ref: https://r-spatial.github.io/stars/articles/stars5.html#reprojecting-a-raster
value_krig %>% st_warp(new_grd) -> value_krig_pred
value_idw["var1.pred"]  %>% st_warp(new_grd) -> value_idw_pred

# Let's convert it to a dataframe now
value_krig_pred <- as.data.frame(value_krig_pred) %>% drop_na()
value_idw_pred <- as.data.frame(value_idw_pred) %>% drop_na()

autoplot.OpenStreetMap(my_osmmap) +
  geom_raster(data = value_idw_pred, aes(x = x, y = y, fill = var1.pred)) +
  scale_fill_viridis_c(name=input$quantity, option = "viridis", alpha=0.4) +
  geom_contour(data = value_idw_pred, aes(x = x, y = y, z=var1.pred), binwidth = ISObinwidth,
               colour ="white", linewidth = 0.1)  +
  geom_text_contour(data = value_idw_pred, aes(x = x, y = y, z=var1.pred), binwidth = ISObinwidth, colour = "white", size=2) + 
  coord_sf(expand = F) +
  labs(title = site,
       subtitle = paste(detectorname,"detector idw"),
       x="Longitude", y="Latitude") +
  theme(legend.position="right",axis.text.x=element_text(angle = -45, hjust = 0))

 ggsave(paste0("printouts/",input$site,"-",detectorname,"-",input$unitName,"-idw.jpg"), dpi=600)
 #ggsave(paste0("printouts/",site,"-",detector,"-",unitName,"-idw.pdf"), dpi=600)


autoplot.OpenStreetMap(my_osmmap) +
  geom_raster(data = value_krig_pred, aes(x = x, y = y, fill = var1.pred)) +
  scale_fill_viridis_c(name=input$quantity, option = "viridis", alpha=0.4) +
  geom_contour(data = value_krig_pred, aes(x = x, y = y, z=var1.pred), binwidth = ISObinwidth,
               colour ="white", linewidth = 0.1)  +
  geom_text_contour(data = value_krig_pred, aes(x = x, y = y, z=var1.pred), binwidth = ISObinwidth, colour = "white", size=2) + 
  coord_sf(expand = F) +
  labs(title = input$site,
       subtitle = paste(detectorname,"detector krig"),
       x="Longitude", y="Latitude") +
  theme(legend.position="right",axis.text.x=element_text(angle = -45, hjust = 0))


# Type ?ggsave in the "Console" to see all options
 ggsave(paste0("printouts/",input$site,"-",detectorname,"-",input$unitName,"-krig.jpg"), dpi=600)
 if (OPEN_WINDOWS){
  viewer(paste0("printouts/",input$site,"-",detectorname,"-",input$unitName,"-krig.jpg"), height = 10)
 }
 #ggsave(paste0("printouts/",site,"-",detector,"-",unitName,"-krig.pdf"), dpi=600)
 
autoplot.OpenStreetMap(my_osmmap) +
  geom_raster(data = value_krig_pred, aes(x = x, y = y, fill = cov)) +
  scale_fill_viridis_c(name="uncertainty [%]", option = "viridis", alpha=0.4) +
  geom_contour(data = value_krig_pred, aes(x = x, y = y, z=var1.pred), binwidth = ISObinwidth,
               colour ="white", linewidth = 0.1)  +
  geom_text_contour(data = value_krig_pred, aes(x = x, y = y, z=var1.pred), binwidth = ISObinwidth, colour = "white", size=2) + 
  coord_sf(expand = F) +
  labs(title = site,
       subtitle = paste(detectorname,"detector krig"),
       x="Longitude", y="Latitude") +
  theme(legend.position="right",axis.text.x=element_text(angle = -45, hjust = 0)) 
 ggsave(paste0("printouts/",input$site,"-",detectorname,"-",input$unitName,"-krig_unc.jpg"), dpi=600)
```
